---
title: "Area Under Disease Gradient (AUDG) Calculation"
author: "Bhim Chaulagain"
date: 2020-07-08T21:13:14-05:00
categories: ["R"]
tags: ["dplyr", "ggplot", "tidyverse"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

Disease gradient information provides the spatial distribution of the epidemics at distance from the focus. If you are interested in learning how to calculate the area under the disease gradient curve to make comparison of several scenarios or treatments, this would be a good resource.

I am using [Steven Worthington](https://gist.github.com/stevenworthington/)'s ipak function to install and load several packages in R.

```{r}
ipak <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}

# usage
packages <- c("tidyverse", "readxl", "agricolae", "ggpubr", "ggpubr", "magrittr", "reshape", "knitr")
ipak(packages)

data <- read_excel("field observation.xlsx") %>% 
        as.data.frame()
head(data, 10)
audpc_data <- data %>% 
  melt(id=c("Distance", "Treatments")) %>%
  group_by(Treatments, variable) %>% 
  summarize(AUDPC = audpc(value, Distance))

audpc_data %>% 
  spread(variable, AUDPC) %>% 
  arrange(Treatments) %>%       
  kable() 

plot_graph <- ggplot(audpc_data, aes(Treatments, AUDPC))+
  geom_boxplot(aes(fill=factor(Treatments))) +
  ggtitle("AUDPC")+
  xlab("Treatment Applied")+
  ylab("Area Under Disease Gradient")+
  theme_bw(base_size = 14)+
  theme(aspect.ratio=1.25,plot.title = element_text(hjust = 0.5), 
        panel.grid=element_blank())+
  scale_fill_manual(name="Treatments", 
                    breaks = c("T1", "T2", 
                               "T3", "T4", "T5",
                               "T6", "T7"),
                    labels=c("T1=Inoculated control", "T2=CP culling at 3 days", "T3=CP cull 'ring' x 3","T4=CP cull at 1 day","T5=FP of CP at 3 days","T6=CP cull + FP of non-CP","T7=FP of entire plot at 3 days"),
                    values=c("#999999","#E69F00","#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00"))


plot_graph

ggsave(filename = "Boxplot.png", width = , height = , units = "in",
       dpi = 600, limitsize = TRUE)
```

