---
title: "Weather based predictors for predictive modeling"
author: "Bhim Chaulagain"
date: 2020-07-08T21:13:14-05:00
categories: ["R"]
tags: ["weather variables", "lubridate", "dplyr"]
---

I would like to share some of the weather predictors that I calculated for predictive modeling of sugarcane rusts from weather data. I have also created several complex weather variables during my research work but I am going to share some with averages/sums of hourly, weekly, monthly, and also specific time in a day. I will share the ways to calculate more complex weather variables in separate posts. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

```{r}
library(readxl)
library(lubridate)
library(plyr)

Wdata <- read_excel("weatherdata.xlsx")
head (Wdata, 10)
str(Wdata)
Wdata$Period <- as.POSIXct(Wdata$Period, format = "%m/%d/%Y %H: %M: %S")
str(Wdata)

#Create Multiple groups to split the data into Days, Weeks and Month

Wdata$days <- floor_date(Wdata$Period, "day")
Wdata$weeks <- floor_date(Wdata$Period, 'week')
Wdata$month <- floor_date(Wdata$Period, 'month')
Wdata$hours <- floor_date(Wdata$Period, 'hours')
```

Lets calculate the weekly average of the temperature. You can use same process for other weather variables just by changing the column name and sums by changing the function name.

Weekly average of temperature!

```{r}
WeeklyAvg_temp <- aggregate(Wdata$`2m T avg (F)`, by = list(Wdata$weeks), FUN = mean)

WeeklyAvg_temp
```
Lets calculate the weekly average of the temperature from(6 AM-9 PM) of each day.

Use lubridate package for subsetting the data which contains the values from 6 AM to 9 PM. 
```{r}
T6AMTo9PM <- subset(Wdata, hour(hours) >= 6 & hour(hours) <= 21 & days >= '2014-01-01' & days <= '2014-12-31')

```
After subsetting, lets calculate for temperature variable!

```{r}
T6AMTo9PM_WeeklyAvg_temp <- aggregate(T6AMTo9PM$`2m T avg (F)`, by = list(T6AMTo9PM$weeks), FUN = mean)
T6AMTo9PM_WeeklyAvg_temp
```
